---
title: "EDS 241 Assignment 4"
author: "Mia Forsline"
date: "3/11/2022"
output: 
  pdf_document:
    toc: false
    number_sections: false
  html_document:
    theme: flatly
    code_folding: show
    toc: true
    toc_float: true
    number_sections: false 
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---
# Introduction 
We plan to estimate the price elasticity of demand for fresh sardines acros 56 points in 4 European countries. We will use monthly data from 2013 - 2019. 

## Data description 
Variables include: 

- `country`
- `port`: where the fish is landed and sold 
- `year`
- `month`
- `price_euro_kg`: price per kg in Euros 
- `volume_sold_kg`: quantity of sardines sold in kg 
- `wind_m_s`

# Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

packages=c("AER", 
           "car", 
           "cowplot",
           "datasets", 
           "dplyr",
           "estimatr",
           "ggplot2",
           "here",
           "janitor",
           "lubridate", 
           "stargazer", 
           "stringr",
           "tibble",
           "tidyverse",
           "tidyr", 
           "tinytex"
           )

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

options(scipen=999) # not scientific notation
```

## Read in the data 
```{r}
data <- read_csv(here("data", "EU_sardines.csv"))

data_clean <- data %>% 
  clean_names() %>% 
  mutate(year = as.factor(year),
         month = as.factor(month), 
         price_euro_kg = as.numeric(price_euro_kg), 
         volume_sold_kg = as.numeric(volume_sold_kg),
         wind_m_s = as.numeric(wind_m_s)
  )
```

# (a) Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.

```{r}
data_log <- data_clean %>% 
  mutate(volume_log = log(volume_sold_kg),
         price_log = log(price_euro_kg))

mdl <- lm_robust(price_log ~ volume_log, data = data_log)
price_elasticity <- round(mdl$coefficients[[2]], digits = 3) 
```

\noindent The price elasticity of demand for sardines is approximately `r price_elasticity`. 

\noindent Table 1 shows the results of a bivariate regression of log-transformed volume sold and log-transformed price per kg of freshs ardines across 56 ports in 4 European contries from 2013 - 2019. 

```{r , results = 'asis', echo = FALSE}
mdl2 <- lm(price_log ~ volume_log, data = data_log)
mdl_se <- starprep(mdl2, 
                   stat = c("std.error"), 
                   se_type = "HC2", 
                   alpha = 0.05)

stargazer(mdl2,  
          se = mdl_se, 
          type = "latex", 
          ci = FALSE, 
          no.space = TRUE, 
          header = FALSE, 
          omit = c("Constant"),  
          omit.stat = c("adj.rsq","ser", "f"),
          covariate.labels = c("Log(Volume)"),
          dep.var.labels = c("Log(Price)"),
          title = "Sardine volume significantly impacts price in Europe", 
          table.placement = "H"
          )
```

# (b) Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? 

## First-stage regression 
```{r}
fs1 <- lm(formula = price_log ~ wind_m_s, data=data_log)
summary(fs1)

wind_coef <- round(fs1$coefficients[[2]], digits = 2)

#F > 10 BUT it is testing all parameters including the constant
```

\noindent The coefficient on wind speed is approximately `r wind_coef`. In other words, on average, when wind speed increases by 1 m/sec, the log price of of sardines per kg increases by `r wind_coef` in Euros. 

Table 
```{r}

```


# (b) Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic. 

## F-test for non-weak and relevant instruments 
```{r}
f <- linearHypothesis(fs1, c("wind_m_s=0"), white.adjust = "hc2")
f_coef <- round(f$F[2], digits = 2)
```

Wind speed is not a weak instrument because the F-test value of `r f_coef` is above our threshold of 10. 

# (c) Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines? 

## Two-stage least square regresion
```{r}
tsls1 <- ivreg(volume_log ~ price_log | wind_m_s, data = data_log)
price_elasticity2 <- round(tsls1$coefficients[[2]], digits = 2)

summary(tsls1)
#takes cares of 1st and 2nd stage regression
#our elasticity of demand = B1 = -1.3744
#we estimate the demand elasticity (slope of the demand curve) using only the variation of log price that comes from wind speed
# 1 demand curve with multiple supply curves moving around (windspeed only affects supply, not demand)
#accounting for endogeneity bias can significantly impact our estimate of B1
```

## Calculate robust standard errors
- use `starprep()`to calculate OLS SEs
- use `coeftest()` to calculate TSLS SEs 
- present SEs using `stargazer()` 
```{r}
se_ols_fs1 <- starprep(mdl2,fs1, stat = c("std.error"), se_type = "HC2", alpha = 0.05)

se_tsls1 <- coeftest(tsls1, vcov = vcovHC(tsls1, type = "HC2"))[, "Std. Error"]

se_models <- append(se_ols_fs1,list(se_tsls1))
stargazer(mdl2, fs1, tsls1, se = se_models, type="text")
```


\noindent The estimated price elasticity of demand for sardines is approximately `r price_elasticity2`. 

# (d) Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.

```{r}
tsls2 <- ivreg(volume_log ~ price_log + as.factor(country) + as.factor(year) + as.factor(month) | wind_m_s, data = data_log)
price_elasticity2 <- round(tsls1$coefficients[[2]], digits = 2)

summary(tsls2)
```

