---
title: "EDS 241 Assignment 5"
author: "Mia Forsline"
date: "3/18/2022"
output: 
  pdf_document:
    toc: false
    number_sections: false
  html_document:
    theme: flatly
    code_folding: show
    toc: true
    toc_float: true
    number_sections: false 
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---
# Introduction 


# Data description 
- `rprice` = inflation-adjusted sales price of house
- `nearinc` = 1 if the house is located near the incinerator, = 0 otherwise
- `age` = age of the house
- `land` = square footage of the lot
- `area` = square footage of the house
- `rooms` = number of rooms in the house
- `year` indicator = 1978 or 1981 

# Research question
What is the impact of the opening of a garbage incinerator on housing values in North Andover, MA ? 

# Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

packages=c("AER", 
           "car", 
           "cowplot",
           "datasets", 
           "dplyr",
           "estimatr",
           "ggplot2",
           "here",
           "janitor",
           "lubridate", 
           "stargazer", 
           "stringr",
           "tibble",
           "tidyverse",
           "tidyr", 
           "tinytex"
           )

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

options(scipen=999) # not scientific notation
```

# Read in the data 

```{r}

```

