---
title: "EDS 241 Assignment 3"
author: "Mia Forsline"
date: "2/20/2022"
output: 
  pdf_document:
    toc: false
    number_sections: false
  html_document:
    theme: flatly
    code_folding: show
    toc: true
    toc_float: true
    number_sections: false 
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

packages=c("cowplot",
           "datasets", 
           "dplyr",
           "estimatr",
           "here", 
           "ggplot2",
           "janitor",
           "stargazer", 
           "stringr",
           "tibble",
           "tidyverse",
           "tidyr", 
           "tinytex"
           )

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

options(scipen=999) # not scientific notation
```

# Introduction

## Goal: to estimate the causal effect of maternal smoking during pregnancy on infant birth weight using the treatment ignorability assumptions

-   data come from the [[Child & Family Data Archive's]{.ul}](https://www.childandfamilydataarchive.org/cfda/pages/cfda/index.html;jsessionid=0A67B5A8C563D765D595DD10EB38A777) [[National Natality Detail Files]{.ul}](https://www.childandfamilydataarchive.org/cfda/cfda/series/36#:~:text=The%20public%2Duse%20Natality%20Detail,births%20occurring%20in%20each%20state.)
-   the data we will be using is a random sample of all births in Pennsylvania during 1989 - 1991
-   each observation is a mother-infant pair

## Variables

-   Outcome variable: `birthwgt` = birth weight of infant in grams

-   Treatment variable: `tobacco` = indicator for maternal smoking

-   Control variables:

    -   `mage` = mother's age
    -   `meduc` = mother's education
    -   `mblack` = 1 if the mother is Black
    -   `alcohol` = 1 if the mother consumed alcohol during pregnancy
    -   `first` = 1 if this is the mother's first child
    -   `diabete` = 1 if the mother is diabetic
    -   `anemia` = 1 if the mother is anemic

## Note

This homework is a simple examination of these data. More research would be needed to obtain a more definitive assessment of the causal effect of smoking on infant health outcomes. Further, for this homework, you can ignore the adjustments to the standard errors that are necessary to reflect the fact that the propensity score is estimated. Just use heteroskedasticity robust standard errors in R. If you are interested, you can read Imbens and Wooldridge (2009) and Imbens (2014) for discussions of various approaches and issues with standard error estimations in models based on the propensity score

```{r}
data <- read_csv(here("data", "SMOKING_EDS241.csv"))

data <- data %>% 
  mutate(mage = as.numeric(mage), 
         meduc = as.numeric(meduc), 
         mblack = as.factor(mblack), 
         alcohol = as.factor(alcohol), 
         first = as.factor(first), 
         diabete = as.factor(diabete), 
         anemia = as.factor(anemia)
         )
```

# (a) What is the unadjusted mean difference in birth weight of infants with smoking and nonsmoking mothers?

```{r}
#Calculate the unadjusted mean difference by hand 
data_smoke <- subset(data, tobacco == 1)
smoke_mean <- round(mean(data_smoke$birthwgt), digits = 2)

data_no_smoke <- subset(data, tobacco == 0)
no_smoke_mean <- round(mean(data_no_smoke$birthwgt), digits = 2)

diff <- no_smoke_mean - smoke_mean

#Calculate the unadjusted mean difference using a linear regression 

mdl <- lm_robust(birthwgt ~ tobacco, data = data)
```

\noindent On average, mothers who smoked during pregnancy gave birth to babies that weighed `r smoke_mean`g. On average, mothers who did not smoke during pregnancy gave birth to babies that weighed `r no_smoke_mean`g. Thus, the unadjusted mean difference in infant birth weights between smoking and nonsmoking mothers is approximately `r diff`g.

# (a) Under what assumption does this correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight? Provide some simple empirical evidence for or against this assumption

```{r}
mdl2 <- lm_robust(mage ~ tobacco, data = data)
mage_diff <- round(mdl2$coefficients[[2]], digits = 2) * -1
```

\noindent Table 1 shows empirical evidence (a statistically significant `tobacco` coefficient) against the assumption that maternal smoking is randomly assigned among mothers who gave birth in Pennsylvania from 1989 - 1991.

```{r , results = 'asis', echo = FALSE}
mdl3 <- lm(mage ~ tobacco, data = data)

mdl_se <- starprep(mdl3, 
                   stat = c("std.error"), 
                   se_type = "HC2", 
                   alpha = 0.05)

stargazer(mdl3,  
          se = mdl_se, 
          type = "latex", 
          ci = FALSE, 
          no.space = TRUE, 
          header = FALSE, 
          omit = c("Constant"),  
          omit.stat = c("adj.rsq","ser", "f"),
          covariate.labels = c("Maternal Smoking"),
          dep.var.labels = c("Mother's Age (Years)"),
          title = "Maternal smoking predicts mothers' ages", 
          table.placement = "H"
          )
```

\noindent This unadjusted mean value is valid if we assume that maternal smoking is randomly assigned among all mothers who gave birth in Pennsylvania during 1989 - 1991. However, that is an unrealistic assumption because this was not a randomized controlled treatment experiment and maternal smoking is not a treatment that can be randomly assigned. In other words, $D$ is not independent of the potential outcomes $Y(1)$ and $Y(0)$.

\noindent Therefore, it is more likely that there are significant differences in the control variables (such as mother's age) depending on if mothers smoked during pregnancy or not. For example, on average, mothers who smoked during pregnancy are `r mage_diff` years younger than mothers who did not smoke during pregnancy (Table 1; p-value < 0.05)

\noindent This demonstrates that mother's age is correlated with whether mothers smoked or not during pregnancy, so this leads to omitted variables bias if we do not include this covariate. 

# (b) Assume that maternal smoking is randomly assigned conditional on the observable covariates listed above. Estimate the effect of maternal smoking on birth weight using a linear regression. Report the estimated coefficient on tobacco and its standard error

```{r}
mdl4 <- lm_robust(birthwgt ~ tobacco + mage + meduc + mblack + alcohol + first + diabete + anemia, data = data)

mdl4_se <- round(mdl4$std.error[[2]], digits = 2)
```

\noindent Table 2 shows the estimated coefficients and standard error of maternal smoking during pregnancy on infant birth weight (g).

```{r , results = 'asis', echo = FALSE}
mdl5 <- lm(birthwgt ~ tobacco + mage + meduc + mblack + alcohol + first + diabete + anemia, data = data)

tobacco_coef <- round(mdl5$coefficients[[2]], digits = 2)

mdl5_sum <- summary(mdl5)$coefficients[, 2]

mdl_se <- starprep(mdl4, 
                   stat = c("std.error"), 
                   se_type = "HC2", 
                   alpha = 0.05)

stargazer(mdl5,  
          se = mdl_se, 
          type = "latex", 
          ci = FALSE, 
          no.space = TRUE, 
          header = FALSE, 
          omit = c("Constant", 
                   "mage",
                   "meduc", 
                   "mblack", 
                   "alcohol", 
                   "first",
                   "diabete",
                   "anemia"),  
          omit.stat = c("adj.rsq","ser", "f"),
          covariate.labels = c("Maternal Smoking"),
          dep.var.labels = c("Infant Birth Weight (g)"),
          title = "Maternal Smoking During Pregnancy Decreases Infant Birth Weights", 
          table.placement = "H"
          )
```

The estimated coefficient of tobacco is `r tobacco_coef` and the standard error is `r mdl4_se`. 

# (c) Use the exact matching estimator to estimate the effect of maternal smoking on birth weight.

\noindent For simplicity, consider the following covariates in your matching estimator:

-   create a 0-1 indicator for mother's age (=1 if mage\>=34)

-   and a 0-1 indicator for mother's education (1 if meduc\>=16)

-   mother's race (mblack)

-   and alcohol consumption indicator (alcohol).

\noindent These 4 covariates will create 2x2x2x2 = 16 cells.

\noindent Report the estimated average treatment effect of smoking on birthweight using the exact matching estimator and its linear regression analogue (Lecture 6, slides 12-14).

```{r}
#Create the covariate indicators 
data_matching <- data %>% 
  mutate(mage_ind = case_when(
    mage >= 34 ~ 1,
    mage < 34 ~ 0),
    meduc_ind = case_when(
      meduc >= 16 ~ 1,
      meduc < 16 ~ 0),
    g = paste0(mage_ind, meduc_ind, mblack, alcohol) #create all combinations of the 4 covariates in a new variable g in the dataframe 
    )
```

## Exact matching estimator

-   Y = `birthwgt`

-   X = `g`

-   D = `tobacco`

```{r}
TIA_table <- data_matching %>%
  group_by(g, tobacco)%>% 
  summarise(n_obs = n(),
            Y_mean = mean(birthwgt, na.rm = T)) %>% 
  gather(variables, values, n_obs:Y_mean) %>% 
  mutate(variables = paste0(variables,"_",tobacco, sep="")) %>% 
  pivot_wider(id_cols = g, names_from = variables, values_from = values) %>% 
  ungroup() %>%  #Ungroup from X values
  mutate(Y_diff = Y_mean_1 - Y_mean_0, #calculate Y_diff
         w_ATE = (n_obs_0+n_obs_1)/(sum(n_obs_0)+sum(n_obs_1)),
         w_ATT = n_obs_1/sum(n_obs_1))%>% #calculate weights
  mutate_if(is.numeric, round, 2) #Round data

stargazer(TIA_table, type= "text", summary = FALSE, digits = 2)
```

## Average treatment effect (ATE) estimation

```{r}
ATE <- round(sum((TIA_table$w_ATE)*(TIA_table$Y_diff)), digits = 2)
```

\noindent The exact matching estimator estimates an average treatment effect (ATE) of `r ATE`.

## Linear regression analogue 

```{r}
mdl6 <- lm_robust(birthwgt ~ tobacco + as.factor(g), data = data_matching)

tobacco_coef <- round(mdl6$coefficients[[2]], digits = 2)
```

\noindent Table 3 shows the estimated coefficients and standard error of maternal smoking during pregnancy on infant birth weight (g). 

```{r , results = 'asis', echo = FALSE}
mdl7 <- lm(birthwgt ~ tobacco + as.factor(g), data = data_matching)

se_models <- starprep(mdl6, stat = c("std.error"), se_type = "HC2", alpha = 0.05)

stargazer(mdl7,  
          se = se_models, 
          type = "latex", 
          ci = FALSE, 
          no.space = TRUE, 
          header = FALSE, 
          omit = c("Constant", 
                   "g"),  
          omit.stat = c("adj.rsq","ser", "f"),
          covariate.labels = c("Maternal Smoking"),
          dep.var.labels = c("Infant Birth Weight (g)"),
          title = "Maternal Smoking During Pregnancy Decreases Infant Birth Weights", 
          table.placement = "H"
          )
```

\noindent The linear regression analogue estimates an average treatment effect of `r tobacco_coef`. 

# (d) Estimate the propensity score for maternal smoking using a logit estimator and based on the following specification: mother's age, mother's age squared, mother's education, and indicators for mother's race, and alcohol consumption.

```{r}
#Create dataframe that includes mother's age squared and necessary indicators 
data_p <- data_matching %>% 
  mutate(tobacco = as.integer(tobacco),
         mage_ind = as.integer(mage_ind), 
         meduc_ind = as.integer(meduc_ind), 
         mage_sq = as.integer(mage ** 2))

#create propensity score model 
ps_model <- glm(tobacco ~ mage_ind + mage_sq + meduc_ind + mblack + alcohol, 
                family = binomial(), 
                data = data_p)
summary(ps_model)

#use estimated logistic equation to create EPS (estimated propensity score)
eps <- predict(ps_model, type = "response") 

#create propensity score weights 
ps_weight <- (data_p$tobacco/eps) + ((1 - data_p$tobacco) / (1 - eps)) 
```

# (e) Use the propensity score weighted regression (WLS) to estimate the effect of maternal smoking on birth weight (Lecture 7, slide 12).

```{r}
wls1 <- lm_robust(formula = birthwgt ~ tobacco, data=data_p, weights=ps_weight)

wls2 <- lm_robust(formula = birthwgt ~ tobacco + mage_ind + mage_sq + meduc_ind + mblack + alcohol, data=data_p, weights=ps_weight)

tobacco_coef_weighted1 <- round(wls1$coefficients[[2]], digits = 2)
tobacco_coef_weighted2 <- round(wls2$coefficients[[2]], digits = 2)
```

```{r}
huxtable::huxreg(wls1, wls2)
```

