---
title: "EDS 241 Assignment 2"
author: "Mia Forsline"
date: "2/4/2022"
output: 
  pdf_document:
    toc: false
    number_sections: yes
  html_document:
    theme: flatly
    code_folding: show
    toc: true
    toc_float: true
    number_sections: false 
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

# Set Up  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

packages=c("cowplot",
           "datasets", 
           "dplyr",
           "estimatr",
           "here", 
           "ggplot2",
           "janitor",
           "stargazer", 
           "stringr",
           "tibble",
           "tidyverse",
           "tidyr", 
           "tinytex"
           )

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

options(scipen=999) # not scientific notation
```


# Read in data 
Variables: 

- `fips`: FIPS code identifying each county
- `NPB`: 1 indicates if the county was regulated under the NO~x~ Budget Program 
- `PctBlack`: fraction of the county population that is African American
- `Dnox_masstons`: change in NO~x~ emissions from all power plants in a county between 2000 - 2008 in tons 
  - if `Dnox_masstons` < 0, then the NO~x~ emissions from power plants declined in that county 

```{r}
data <- readxl::read_excel(here("data", "NBP.xls"))

data_clean <- data %>% 
  janitor::clean_names()
```

# (a) Make a histogram depicting the distribution of Dnox_masstons.

```{r}
nox_lab <- expression(paste("Change in NOx"["3"]," emissions from 2000 - 2008 (tons)"))

hist <- ggplot(data = data_clean, aes(x = dnox_masstons)) + 
  geom_histogram() + 
  theme_classic() + 
  labs(x = nox_lab, 
       y = "Frequency")
```


\newpage
<center>**Figure 1: Histogram of change in NO~x~ emissions from all power plants in a county between 2000 and 2008 (in tons)**</center>

```{r , fig.width = 4, fig.height = 3, eval=TRUE, echo=FALSE}
hist
```

\noindent Figure 1 shows the skewed distribution of change of NO~x~ emissions from 2000 - 2008. There is a long tail to the left, and most of the data centers around zero. Each data point represents one county in California (n = 485). 

# (b) Create an indicator =1 if the county has PctBlack above the sample median, and =0 otherwise (in the rest of the assignment, I refer to this variable as ‘D’). What is the average of PctBlack for counties above the median, i.e. counties for which D=1?

```{r}
median <- round(median(data_clean$pct_black), digits = 2)

data_d <- data_clean %>% 
  mutate(D = case_when(
    pct_black > median ~ 1,
    pct_black <= median ~ 0), 
    D = as.factor(D), 
    nbp = as.factor(nbp)
    )

data_above_median <- subset(data_d, D == 1)
avg_pct_black <- round(mean(data_above_median$pct_black), digits = 2) 
```

\noindent For counties with a percentage of Black residents above the median (n = 240), the average percentage of Black individuals is `r avg_pct_black`%. 

# (c) Estimate a regression of Dnox_masstons on NBP. Interpret the estimated intercept and the coefficient on NBP.

\noindent To more formally analyze the relationship between the change in NO~x~ emissions from all power plants in a county between 2000 - 2008 in tons and if a county was regulated under the NO~x~ Budget Program or not, we estimate the following regression: 

\begin{align}
  Y_{i} = \beta_0 + \beta_1 X_{1i} + u_{i}
\end{align}

\noindent where $Y_{i}$ is the tons of NO~x~ emissions from county $i$, $X_{1i}$ is the dummy variable indicating if a county was regulated under the NO~x~ Budget Program or not, and $u_{i}$ the regression error term. 

```{r}
mod <- lm_robust(dnox_masstons ~ nbp, data = data_d)
#summary(mod)
abs_value_intercept <- round(mod$coefficients[[1]] * -1, digits = 2)
npb <- round(mod$coefficients[[2]] * -1, digits = 2)
```

\noindent The intercept value tells us that, on average, counties that did not participate in the NO~x~ Budget Program reported a `r abs_value_intercept` tons decrease in the change of NO~x~ emissions from all power plants in the county from 2000 to 2008. 

\noindent The `npb` coefficient tells us that, on average, counties that participated in the NO~x~ Budget Program reported a `r npb` tons decrease in the change of NO~x~ emissions from all power plants in the county from 2000 to 2008 compared to counties that did not participate in the NO~x~ Budget Program. 

# (d) Create an interaction between the variables NBP and D. Estimate a regression of Dnox_masstons on NBP, D, and this interaction. Interpret each estimated regression coefficient, including the intercept.

```{r}
mod2 <- lm_robust(dnox_masstons ~ nbp + D + nbp:D, data = data_d)
#summary(mod2)
abs_value_intercept <- round(mod2$coefficients[[1]] * -1, digits = 2)
npb_coefficient <- round(mod2$coefficients[[2]] * -1, digits = 2)
D_coefficient <- round(mod2$coefficients[[3]] * -1, digits = 2)
interaction <- round(mod2$coefficients[[4]], digits = 2)
```

\noindent Table 1 shows the estimated coefficients from estimating Model 1 (`dnox_masstons ~ nbp + D`) and Model 2 (`dnox_masstons ~ nbp + D + nbp:D`). 

```{r , results = 'asis', echo = FALSE}

mod <- lm(dnox_masstons ~ nbp, data = data_d)
mod2 <- lm(dnox_masstons ~ nbp + D + nbp:D, data = data_d)

robust_se <- starprep(mod, mod2) #display heteroskedasticity-robust standard errors in the table 

stargazer(mod, mod2, type = "latex", 
          ci=FALSE, 
          no.space = TRUE, 
          header = FALSE, 
          omit.stat = c("adj.rsq","ser", "f"),
          covariate.labels = c("Constant", "NBP", "D", "NPB:D"), 
          dep.var.labels = c("NOx Change (tons)"),
          dep.var.caption = c(""),
          title = "Model Coefficients", 
          table.placement = "H")
```


\noindent The intercept value tells us that, on average, counties that *did not participate* in the NO~x~ Budget Program and have a fraction of the county population that is African American *below* the median value reported a `r abs_value_intercept` tons decrease in the change of NO~x~ emissions from all power plants in the county from 2000 to 2008. 

\noindent The `nbp` coefficient tells us that, for all counties that have a fraction of the county population that is African American *below* the median value, on average, counties that *participated* in the NO~x~ Budget Program reported a `r npb_coefficient` tons decrease in NO~x~ emissions compared to counties that *did not participate* in the NO~x~ Budget Program. 

\noindent The `D` coefficient tells us that, for all counties that *did not participate* in the NO~x~ Budget Program, on average, counties that have a fraction of the county population that is African American *above* the median value reported `r D_coefficient` tons decrease in the change NO~x~ emissions compared to counties that have a fraction of the county population that is African American *below* the median value. 

\noindent The interaction tells us that, on average that: 

- a) counties that *participated* in the NO~x~ Budget Program, regardless of their fraction of the county population that is African American
- b) counties that have a fraction of the county population that is African American *above* the median reported value of `r median` %, regardless of whether they participated in the NO~x~ Budget Program 
- c) and counties that *participated* in the NO~x~ Budget Program and have a fraction of the county population that is African American *above* the median value reported `r interaction` more tons of NO~x~ emissions from all power plants from 2000 to 2008. 

# (e) What is the predicted Dnox_masstons in a county that was not regulated under NBP and where PctBlack is above the sample median (i.e., where D=1)? Report the 95% confidence interval for this prediction. Make sure to use “heteroskedasticity-robust” standard errors.

```{r}
mod2 <- lm_robust(dnox_masstons ~ nbp + D + nbp:D, data = data_d)

pred_constraints=data.frame(nbp=c("0"), D=c("1"))

pred <- predict(mod2, newdata=pred_constraints, se.fit=TRUE, interval='confidence')
pred_value <- round(pred$fit[[1]] * -1, digits = 2)
lower <- round(pred$fit[[2]], digits = 2)
upper <- round(pred$fit[[3]], digits = 2)
```

\noindent A county that did not participate in the NO~x~ Budget Program and has a fraction of the county population that is African American above the median is predicted to have emitted approximately `r pred_value` tons less from 2000 to 2008. 

\noindent The 95% confidence interval is bounded by a lower bount of `r lower` and an upper bound of `r upper`. 
