---
title: "EDS 241 Assignment 2"
author: "Mia Forsline"
date: "2/4/2022"
output: 
  html_document:
    theme: flatly
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

# Set Up 

```{r}
library(tidyverse)
library(here)
```

# Read in data 
- `fips`: FIPS code identifying each county
- `NPB`: 1 indicates if the county was regulated under the NO~x~ Budget Program 
- `PctBlack`: fraction of the county population that is African American
- `Dnox_masstons`: change in annual NO~x~ emissions from all power plants in a county between 2000 - 2008 in tons 
  - if `Dnox_masstons` < 0, then the NO~x~ emissions from power plants declined in that county 
```{r}
data <- readxl::read_excel(here("data", "NBP.xls"))
```

# (a) Make a histogram depicting the distribution of Dnox_masstons.

# (b) Create an indicator =1 if the county has PctBlack above the sample median, and =0 otherwise (in the rest of the assignment, I refer to this variable as ‘D’). What is the average of PctBlack for counties above the median, i.e. counties for which D=1?

# (c) Estimate a regression of Dnox_masstons on NBP. Interpret the estimated intercept and the coefficient on NBP.

# (d) Create an interaction between the variables NBP and D. Estimate a regression of Dnox_masstons on NBP, D, and this interaction. Interpret each estimated regression coefficient, including the intercept.

# (e) What is the predicted Dnox_masstons in a county that was not regulated under NBP and where PctBlack is above the sample median (i.e., where D=1)? Report the 95% confidence interval for this prediction. Make sure to use “heteroskedasticity-robust” standard errors.

